\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

\subsubsection{Notation and organization}
We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. The set of natural numbers is denoted as $\mathbb{N}$. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. The symbol $|\cdot|$ denotes the absolute value, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
\begin{align}
\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
\label{dPhidt}	
\end{align}
wherein the functional derivative
\begin{align}
\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
\label{FunctionalDerivative}
\end{align}
Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
\begin{strip}
\hrule
\begin{subequations}
\begin{align}
\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
\end{pmatrix}, \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
\end{align}
\label{ExplicitdPhidt}	
\end{subequations}
\hrule	
\end{strip}
where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.

>>>>

\subsubsection{Notation and organization}
We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. The set of natural numbers is denoted as $\mathbb{N}$. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
\begin{align}
\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
\label{dPhidt}	
\end{align}
wherein the functional derivative
\begin{align}
\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
\label{FunctionalDerivative}
\end{align}
Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
\begin{strip}
\hrule
\begin{subequations}
\begin{align}
\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
\end{pmatrix}, \begin{pmatrix}
 \bm{\eta}\\
 -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
 \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
\end{align}
\label{ExplicitdPhidt}	
\end{subequations}
\hrule	
\end{strip}
where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
